Unifying Lower Bounds on Prediction Dimension of Consistent Convex Surrogates

### 2021-01-26 colt-21/cvx-flats.tex
Still need to explain why spn \simplexp = dim affhull \simplexp

\begin{theorem}\label{thm:consistent-implies-indir-elic}
	For a convex surrogate $L \in \L$, if the pair $(L, \psi)$ is $\D$-consistent with respect to a property $\gamma$ or a loss $\ell$ eliciting $\gamma$, then $(L, \psi)$ indirectly elicits $\gamma$.\jessiet{How do we want to handle consistency here- we want it to be specific to the distribution}\jessie{Require $\P \supseteq \cup_{D \in \D}\cup_x D_x$ and convex?  With our assumptions on the loss (namely, the expected loss is finite everywhere for all $p$), we should be able to apply Fubini-Tonelli, so that $\P \supseteq \cup_D \cup_x D_x$ and convex is sufficient, as any $D_X$ will be covered by this since the joint expectation is the same as $\E_X \E_Y L(u;D_X)$}\jessiet{Need $\X$ and $\Y$ to be $\sigma$-finite, which I think is immediate from them being probability measures} \jessie{Union of convex hulls of $D_x$ is convex, and use the fact that we .  Better : assume $\P = \{D_x : x \in \X and D \in \D\}$}
\end{theorem}
\begin{proof}
%  \raft{Maybe ref the consistency definition here explicitly, at the top and the end}
	By Lemma~\ref{lem:consistent-loss-implies-prop}, it suffices to show the result for consistency with respect to a property $\gamma$, setting $\gamma := \prop{\ell}$ if $\ell$ is given instead.
	We show the contrapositive; suppose $(L, \psi)$ does not indirectly elicit $\gamma$, meaning we have some $p \in \P$ so that $u \in \Gamma(p)$ but $\psi(u) \not \in \gamma(p)$, where $\Gamma := \prop{L}$.
	Observe that we use the fact $\Gamma(p) \neq \emptyset$.
	Consider the constant sequence $\{f_m\}$ with $f_m(x) = u$ for all $m,x$, and take $D$ with full support on some $x \in \X$, and let $D_x = p$.
  Since $u \in \Gamma(p)$, we observe $\E_D L(f_m(X), Y) = \inf_f L(f(X),Y)$ for all $m$; in particular $\E_D L(f_m(X), Y) \to \inf_f L(f(X),Y)$.
%  \raft{Maybe clearer to first say $\E_D L(f_m(X), Y) = \inf_f L(f(X),Y)$ for all $m$, and then the easy implication of the convergence \jessie{Done}}
	However, we have $\E_X \propdis(\psi \circ f_m(X), D_X) = \propdis(\psi(u_m), p) = \propdis(\psi(u), p) \neq 0$, since $\psi(u) \not \in \gamma(p)$.
%  \raft{Almost -- just clarify that this is the expectation over $D$ too by construction \jessie{Kinda done?}}
	Thus, we observe $(L, \psi)$ is not consistent with respect to $\gamma$ (Definition~\ref{def:consistent-prop}).
\end{proof}

Comment about how the assumption on sets isn't a big deal.

@@

\jessie{Here, we need to mention $\D$-consistency in the ``in particular''s.  And specify $\P \subseteq \cup_{D \in \D} \cup_x D_x$ or whatever condition we settle on.}
\begin{corollary}\label{cor:Pcodim-flat-single-val-prop}
Let target property $\gamma:\P \toto \R$ and $d\in\mathbb N$ be given.
Let $p \in \P$ with $|\gamma(p)| = 1$, and take $\gamma(p) = \{r\}$.
If there is no $d$-flat $F$ with $p \in F \subseteq \gamma_r$, then no $L\in\L_d$ indirectly elicits $\gamma$\raft{To discuss next meeting: maybe it's better to say $\eliccvx(\gamma) > d$ instead of the indirectly elicits bit.  Both here and the next Cor.}, and in particular, there is no convex surrogate $L \in \L_d$ consistent with respect to $\gamma$.
%  %commented out 20 Jan 2021 8:37am - Jessie.  switching first statement to be the contrapositive
%  Suppose we are given a target property $\gamma:\P \toto \R$, where $\P\subseteq\simplex$, indirectly elicited by a convex loss $L \in \L_d$.
%  For all $p \in \P$ with $|\gamma(p)| = 1$, and $\gamma(p) = \{r\}$, there is some flat $F$ of \Pcodimension at most $d$ such that $p \in F \subseteq \gamma_r$.
%  In particular, if there is a $p \in \P$ such that $\gamma(p) = \{r\}$ such that there is no flat $F$ of \Pcodimension at most $d$ such that $p \in F \subseteq \gamma_r$, then there is no convex surrogate $L \in \L_d$ consistent with respect to $\gamma$.
\end{corollary}
\begin{proof}
	Let $(L, \psi)$ indirectly elicit $\gamma$ and the convex function $L$ and elicit $\Gamma$.
	As $\Gamma$ is non-empty, there is some $u \in \Gamma(p)$.
	Since $\gamma$ is single-valued at $p$, we have $r = \psi(u)$; by Lemma~\ref{thm:convex-flats-inf-dim}, we know there is a flat $F = \zeros{V_{u,p}}$ so that $p \in F \subseteq \Gamma_u$.
	Moreover, by construction, we know it is a $d$-flat.
	By definition of indirect elicitation, we additionally have $\Gamma_u \subseteq \gamma_r$.
	Thus, we have $p \in F \subseteq \gamma_r$.

	Now, if we do not have a flat satisfying the above conditions, then we do not have any $L$ so that $\prop{L}$ indirectly elicits $\gamma$, and in turn do not have a link such that $L$ is consistent with respect to $\gamma$ by Theorem~\ref{thm:consistent-implies-indir-elic}.
\end{proof}

@@

\begin{corollary}\label{cor:Pcodim-flat-elic-relint-prop} 
  Let an elicitable target property $\gamma:\P \toto \R$ be given, where $\P\subseteq\simplex$ is defined over a finite set of outcomes $\Y$, and let $d\in\mathbb N$.
Let $p \in \relint{\P}$.
If there is no $d$-flat $F$ with $p \in F \subseteq \gamma_r$, then no $L\in\L_d$ indirectly elicits $\gamma$, and in particular, there is no convex surrogate $L \in \L_d$ consistent with respect to $\gamma$.
%
%  %commented out 20 Jan 2021 at 8:40a - Jessie - switching first statement to be the contrapositive.
%  Suppose we are given a target property $\gamma:\P \toto \R$, where $\P\subseteq\simplex$ for a finite set of outcomes $\Y$, indirectly elicited by a convex loss $L \in \L_d$.
%	For all $p \in \relint{\P}$, there is some flat $F$ of \Pcodimension at most $d$ such that $p \in F \subseteq \gamma_r$.
%	Moreover, if there is some $p \in \relint{\P}$ so that there is not a flat $F \subseteq \gamma_r$ of \Pcodimension at most $d$ then there is no $d$-dimensional convex surrogate $L$ consistent with respect to $\gamma$.
\end{corollary}
\begin{proof}
	Let $(L, \psi)$ indirectly elicit $\gamma$ and the convex function $L$ and elicit $\Gamma$.
	As $\Gamma$ is non-empty, there is some $u \in \Gamma(p)$, and suppose $r' = \psi(u)$.
	Take $F \subseteq \Gamma_u$ to be the flat that exists by Lemma~\ref{thm:convex-flats-inf-dim}.
	If $r = r'$, then $p \in F \subseteq \Gamma_u \subseteq \gamma_r$ by indirect elicitation.
	Otherwise, by Lemma~\ref{lem:set-valued-prop-flats}, for elicitable properties with $p \in \gamma_r \cap \gamma_{r'}$, we observe $p \in F\subseteq \gamma_r \iff p \in F \subseteq \gamma_{r'}$.

	As above, if we do not have a flat satisfying the above conditions, then we do not have any $L$ so that $\prop{L}$ indirectly elicits $\gamma$, and in turn do not have a link such that $L \in \L_d$ is consistent with respect to $\gamma$ by Theorem~\ref{thm:consistent-implies-indir-elic}.
\end{proof}

@@

However, consider any $1$-flat containing $p$ (the uniform distribution).
When intersected with the simplex, one can see that any line in the simplex through $p$ also leaves the cell $\gamma_\bot$, which contains $p$.
See Figure~\ref{fig:flats-bound} for intuition; a $1$-flat would be a line in such a figure. \jessiet{Justify why}
Therefore, we have no $1$-flat containing $p$ staying in $\gamma_\bot$, so we obtain a better lower bound, $\ccdim(\ell^{abs}) \geq 2$.

@@

Throughout, given a point (typically a distribution) $p$ and convex set $P$, we define $P_p := P - \{p\}$.
\jessiet{Get rid of this notation?}

\begin{lemma}\label{lem:finite-relint-dim}
  Let the $d$-flat $F\subseteq \P$ (defined over finite $\Y$) contain some $p\in\relint{\P}$.
  Then 
  \begin{enumerate}
  	\item[(i)] $p \in \relint{F}$; 
  	\item[(ii)] $\Scr_F(p) = \spn(F_p)$ and $\dim(\Scr_F(p)) \geq \dim(\spn(\P_p)) - d$.
  \end{enumerate}
\end{lemma}
\jessiet{Bo had a good point that $\spn(F_p)$ is probably easier to understand, instead of having an extra layer of notation in there.  Do we want to stick with $\Scr_F(p)$ or $\spn(F_p)$?}
\begin{proof}
  %\raf{Moved; see margin comment:  for the function $W : \Y \to \reals^d$ such that $F = \zeros{W}$.  (Such a function must exist by $F$ being a $d$-flat.)}

  As $F$ is a $d$-flat, we have some $W:\Y \to \reals^d$ such that $F = \zeros{W}$.
  Define $T_W:\spn(\P_p)\to\reals^d, v\mapsto \E_v W$.

>>>>

In what follows, we assume $\P = \{D_x : x \in \X, D \in \D\}$ and is a convex set.
\begin{theorem}\label{thm:consistent-implies-indir-elic}
	For a convex surrogate $L \in \L$, if the pair $(L, \psi)$ is $\D$-consistent with respect to a property $\gamma: \P \toto \R$ or a loss $\ell$ eliciting $\gamma$, then $(L, \psi)$ indirectly elicits $\gamma$.\jessiet{How do we want to handle consistency here- we want it to be specific to the distribution}\jessie{Require $\P \supseteq \cup_{D \in \D}\cup_x D_x$ and convex?  With our assumptions on the loss (namely, the expected loss is finite everywhere for all $p$), we should be able to apply Fubini-Tonelli, so that $\P \supseteq \cup_D \cup_x D_x$ and convex is sufficient, as any $D_X$ will be covered by this since the joint expectation is the same as $\E_X \E_Y L(u;D_X)$}\jessiet{Need $\X$ and $\Y$ to be $\sigma$-finite, which I think is immediate from them being probability measures} 
	%\jessie{Union of convex hulls of $D_x$ is convex, and use the fact that we .  Better : assume $\P = \{D_x : x \in \X and D \in \D\}$}
\end{theorem}
\begin{proof}
%  \raft{Maybe ref the consistency definition here explicitly, at the top and the end}
	By Lemma~\ref{lem:consistent-loss-implies-prop}, it suffices to show the result for consistency with respect to a property $\gamma$, setting $\gamma := \prop{\ell}$ if $\ell$ is given instead.
	We show the contrapositive; suppose $(L, \psi)$ does not indirectly elicit $\gamma$, meaning we have some $p \in \P$ so that $u \in \Gamma(p)$ but $\psi(u) \not \in \gamma(p)$, where $\Gamma := \prop{L}$.
	Observe that we use the fact $\Gamma(p) \neq \emptyset$.
	Consider the constant sequence $\{f_m\}$ with $f_m(x) = u$ for all $m,x$, and take $D$ with full support on some $x \in \X$, and let $D_x = p$.
  Since $u \in \Gamma(p)$, we observe $\E_D L(f_m(X), Y) = \inf_f L(f(X),Y)$ for all $m$; in particular $\E_D L(f_m(X), Y) \to \inf_f L(f(X),Y)$.
%  \raft{Maybe clearer to first say $\E_D L(f_m(X), Y) = \inf_f L(f(X),Y)$ for all $m$, and then the easy implication of the convergence \jessie{Done}}
	However, we have $\E_X \propdis(\psi \circ f_m(X), D_X) = \propdis(\psi(u_m), p) = \propdis(\psi(u), p) \neq 0$, since $\psi(u) \not \in \gamma(p)$.
%  \raft{Almost -- just clarify that this is the expectation over $D$ too by construction \jessie{Kinda done?}}
	Thus, we observe $(L, \psi)$ is not consistent with respect to $\gamma$ (Definition~\ref{def:consistent-prop}).
\end{proof}

\jessie{Comment about how the assumption on sets isn't a big deal.}
While the assumption $\P = \{D_x : x \in \X, D \in \D\}$ may seem restrictive, it is primarily made for the sake of bookkeeping.
In particular, when $\Y$ is a finite set, we often take $\P = \simplex$, in which case, the assumption is unnecessary.
The primary reason we desire $\P$ more general than $\simplex$ is when there are infinitely many possible outcomes; for example, squared loss only elicits the expected value on distributions whose variance is finite.
In this case, we still do not find the assumption very restrictive, and find it covers most cases one might practically be interested in.

@@

\jessie{Here, we need to mention $\D$-consistency in the ``in particular''s.  And specify $\P = \{D_x : x \in \X, D \in \D\}$ or whatever condition we settle on.}
\begin{corollary}\label{cor:Pcodim-flat-single-val-prop}
Let target property $\gamma:\P \toto \R$ and $d\in\mathbb N$ be given.
Let $p \in \P$ with $|\gamma(p)| = 1$, and take $\gamma(p) = \{r\}$.
If there is no $d$-flat $F$ with $p \in F \subseteq \gamma_r$, then no $L\in\L_d$ indirectly elicits $\gamma$\raft{To discuss next meeting: maybe it's better to say $\eliccvx(\gamma) > d$ instead of the indirectly elicits bit.  Both here and the next Cor.}, and in particular, there is no convex surrogate $L \in \L_d$ that is  $\D$-consistent with respect to $\gamma$.
%  %commented out 20 Jan 2021 8:37am - Jessie.  switching first statement to be the contrapositive
%  Suppose we are given a target property $\gamma:\P \toto \R$, where $\P\subseteq\simplex$, indirectly elicited by a convex loss $L \in \L_d$.
%  For all $p \in \P$ with $|\gamma(p)| = 1$, and $\gamma(p) = \{r\}$, there is some flat $F$ of \Pcodimension at most $d$ such that $p \in F \subseteq \gamma_r$.
%  In particular, if there is a $p \in \P$ such that $\gamma(p) = \{r\}$ such that there is no flat $F$ of \Pcodimension at most $d$ such that $p \in F \subseteq \gamma_r$, then there is no convex surrogate $L \in \L_d$ consistent with respect to $\gamma$.
\end{corollary}
\begin{proof}
	Let $(L, \psi)$ indirectly elicit $\gamma$ and the convex function $L$ and elicit $\Gamma$.
	As $\Gamma$ is non-empty, there is some $u \in \Gamma(p)$.
	Since $\gamma$ is single-valued at $p$, we have $r = \psi(u)$; by Lemma~\ref{thm:convex-flats-inf-dim}, we know there is a flat $F = \zeros{V_{u,p}}$ so that $p \in F \subseteq \Gamma_u$.
	Moreover, by construction, we know it is a $d$-flat.
	By definition of indirect elicitation, we additionally have $\Gamma_u \subseteq \gamma_r$.
	Thus, we have $p \in F \subseteq \gamma_r$.

	Now, if we do not have a flat satisfying the above conditions, then we do not have any $L$ so that $\prop{L}$ indirectly elicits $\gamma$, and in turn do not have a link such that $L$ is $\D$-consistent with respect to $\gamma$ by Theorem~\ref{thm:consistent-implies-indir-elic}.
\end{proof}

@@

\begin{corollary}\label{cor:Pcodim-flat-elic-relint-prop} 
  Let an elicitable target property $\gamma:\P \toto \R$ be given, where $\P\subseteq\simplex$ is defined over a finite set of outcomes $\Y$, and let $d\in\mathbb N$.
Let $p \in \relint{\P}$.
If there is no $d$-flat $F$ with $p \in F \subseteq \gamma_r$, then no $L\in\L_d$ indirectly elicits $\gamma$, and in particular, there is no convex surrogate $L \in \L_d$ that is $\D$-consistent with respect to $\gamma$.
%
%  %commented out 20 Jan 2021 at 8:40a - Jessie - switching first statement to be the contrapositive.
%  Suppose we are given a target property $\gamma:\P \toto \R$, where $\P\subseteq\simplex$ for a finite set of outcomes $\Y$, indirectly elicited by a convex loss $L \in \L_d$.
%	For all $p \in \relint{\P}$, there is some flat $F$ of \Pcodimension at most $d$ such that $p \in F \subseteq \gamma_r$.
%	Moreover, if there is some $p \in \relint{\P}$ so that there is not a flat $F \subseteq \gamma_r$ of \Pcodimension at most $d$ then there is no $d$-dimensional convex surrogate $L$ consistent with respect to $\gamma$.
\end{corollary}
\begin{proof}
	Let $(L, \psi)$ indirectly elicit $\gamma$ and the convex function $L$ and elicit $\Gamma$.
	As $\Gamma$ is non-empty, there is some $u \in \Gamma(p)$, and suppose $r' = \psi(u)$.
	Take $F \subseteq \Gamma_u$ to be the flat that exists by Lemma~\ref{thm:convex-flats-inf-dim}.
	If $r = r'$, then $p \in F \subseteq \Gamma_u \subseteq \gamma_r$ by indirect elicitation.
	Otherwise, by Lemma~\ref{lem:set-valued-prop-flats}, for elicitable properties with $p \in \gamma_r \cap \gamma_{r'}$, we observe $p \in F\subseteq \gamma_r \iff p \in F \subseteq \gamma_{r'}$.

	As above, if we do not have a flat satisfying the above conditions, then we do not have any $L$ so that $\prop{L}$ indirectly elicits $\gamma$, and in turn do not have a link such that $L \in \L_d$ is $\D$-consistent with respect to $\gamma$ by Theorem~\ref{thm:consistent-implies-indir-elic}.
\end{proof}

@@

However, consider any $1$-flat containing $p$ (the uniform distribution).
When intersected with the simplex, one can see that any line in the simplex through $p$ also leaves the cell $\gamma_\bot$, which contains $p$.
See Figure~\ref{fig:flats-bound} for intuition; a $1$-flat through $p \in \relint{\simplex}$ would be a line in such a figure.
Therefore, we have no $1$-flat containing $p$ staying in $\gamma_\bot$, so we obtain a better lower bound, $\ccdim(\ell^{abs}) \geq 2$.

@@

\begin{lemma}\label{lem:finite-relint-dim}
  Let the $d$-flat $F\subseteq \P$ (defined over finite $\Y$) contain some $p\in\relint{\P}$.
  Then 
  \begin{enumerate}
  	\item[(i)] $p \in \relint{F}$; 
  	\item[(ii)] $\dim(\Scr_F(p)) \geq \dim(\spn(\P - \{p\})) - d$.
  \end{enumerate}
\end{lemma}
\jessiet{Bo had a good point that $\spn(F_p)$ is probably easier to understand, instead of having an extra layer of notation in there.  Do we want to stick with $\Scr_F(p)$ or $\spn(F_p)$?}
\begin{proof}
  %\raf{Moved; see margin comment:  for the function $W : \Y \to \reals^d$ such that $F = \zeros{W}$.  (Such a function must exist by $F$ being a $d$-flat.)}

  As $F$ is a $d$-flat, we have some $W:\Y \to \reals^d$ such that $F = \zeros{W}$.
  Throughout, given a point (typically a distribution) $p$ and convex set $P$, we define $P_p := P - \{p\}$.
  Define $T_W:\spn(\P_p)\to\reals^d, v\mapsto \E_v W$.
